### 2024-08-10 machine-learning.qmd
Built site for gh-pages

seasonalData <- powerjoin::power_full_join(
  playerSeasonMerged,
  playerMerged,
  by = "gsis_id",
  conflict = coalesce_xy
)

seasonalAndWeeklyData <- powerjoin::power_full_join(
  playerSeasonWeekMerged,
  seasonalData,
  by = c("gsis_id","season"),
  conflict = coalesce_xy
)

@@

seasonalData %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()
```

@@

seasonalData_subset <- seasonalData %>% 
  dplyr::select(-any_of(c(
    "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
    "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "fantasy_data_id", "sleeper_id", "pfr_id",
    "pfr_player_id", "cfb_player_id", "pfr_player_name", "esb_id", "gsis_it_id", "smart_id",
    "college", "college_name", "team_abbr", "current_team_id", "college_conference", "draft_club", "status_description_abbr",
    "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
    "football_name", "team"
  ))) %>% 
  filter(!is.na(season))
```

@@

```{r}
seasonalData_subset <- seasonalData_subset %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  mutate(
    fantasyPoints_lag = lead(fantasy_points)
  ) %>% 
  ungroup()
```

@@

- CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
- IMPUTE THE TRAINING AND TEST DATA SEPARATELY

@@

seasonalData_subset_imp <- missForest::missForest(
  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, years_of_experience, position_group, fantasyPoints_lag) %>% data.frame,
  verbose = TRUE,
  mtry = 100,
  parallelize = "forests")
```

@@

```{r}
# Specify data to impute
dataToImpute <- seasonalData_subset %>% 
  select(gsis_id, season, ageCentered20, ageCentered20Quadratic, fantasy_points, years_of_experience) %>% 
  filter(!is.na(gsis_id), !is.na(season)) %>% 
  mutate(ID = as.integer(as.factor(as.character(gsis_id)))) %>% 
  arrange(gsis_id, season)

@@

```{r}
smerf <- LongituRF::MERF(
  X = seasonalData_subset_imp$ximp[,c("position_group")] %>% as.matrix(),
  Y = seasonalData_subset_imp$ximp[,c("fantasyPoints_lag")],
  Z = seasonalData_subset_imp$ximp[,c("years_of_experience")] %>% as.matrix(),
  id = seasonalData_subset$gsis_id,
  time = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
  mtry = 2,
  ntree = 500,
  sto = "BM")

>>>>

seasonalData <- powerjoin::power_full_join(
  playerSeasonMerged,
  playerMerged,
  by = "gsis_id",
  conflict = coalesce_xy
) %>% 
  filter(!is.na(season))

seasonalAndWeeklyData <- powerjoin::power_full_join(
  playerSeasonWeekMerged,
  seasonalData,
  by = c("gsis_id","season"),
  conflict = coalesce_xy
) %>% 
  filter(!is.na(week))

@@

seasonalData %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()

seasonalAndWeeklyData %>% 
  group_by(gsis_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()
```

```{r}
seasonalData <- seasonalData %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  mutate(
    fantasyPoints_lag = lead(fantasy_points)
  ) %>% 
  ungroup()

seasonalData$fantasyPoints_lag[is.na(seasonalData$fantasyPoints_lag)] <- 0
```

### Fill in Missing Data {#sec-fillMissingData}

```{r}
seasonalData <- seasonalData %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  fill(
    player_name, player_display_name, pos, position, position_group,
    .direction = "downup") %>% 
  ungroup()
```

@@

seasonalData_subset <- seasonalData %>% 
  dplyr::select(-any_of(c(
    "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
    "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "fantasy_data_id", "sleeper_id", "pfr_id",
    "pfr_player_id", "cfb_player_id", "pfr_player_name", "esb_id", "gsis_it_id", "smart_id",
    "college", "college_name", "team_abbr", "current_team_id", "college_conference", "draft_club", "status_description_abbr",
    "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
    "football_name", "team"
  )))
```

@@

### Separate by Position {#sec-separateByPosition}

```{r}
seasonalData_subsetQB <- seasonalData_subset %>% 
  filter(position == "QB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    completions:rushing_2pt_conversions, contains(".pass"))

seasonalData_subsetRB <- seasonalData_subset %>% 
  filter(position == "RB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_subsetWR <- seasonalData_subset %>% 
  filter(position == "WR") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_subsetTE <- seasonalData_subset %>% 
  filter(position == "TE") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))
```

@@

- SEPARATE BY POSITION
- CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
- IMPUTE THE TRAINING AND TEST DATA SEPARATELY

@@

seasonalData_subsetQB_imp <- missForest::missForest(
  seasonalData_subsetQB %>% 
    select(ageCentered20, ageCentered20Quadratic, years_of_experience, passing_epa, pacr, dakota, rushing_epa) %>% 
    data.frame(),
  verbose = TRUE,
  mtry = 100,
  parallelize = "forests")
```

@@

```{r}
# Specify data to impute
dataToImpute <- seasonalData_subset %>% 
  select(gsis_id, season, ageCentered20, ageCentered20Quadratic, fantasyPoints_lag, years_of_experience, passing_epa) %>% 
  filter(!is.na(gsis_id), !is.na(season)) %>% 
  mutate(ID = as.integer(as.factor(as.character(gsis_id)))) %>% 
  arrange(gsis_id, season)

@@

```{r}
smerf <- LongituRF::MERF(
  X = seasonalData_subsetQB_imp$ximp[,c("passing_epa")] %>% as.matrix(),
  Y = seasonalData_subsetQB$fantasyPoints_lag,
  Z = seasonalData_subsetQB_imp$ximp[,c("years_of_experience")] %>% as.matrix(),
  id = seasonalData_subsetQB$gsis_id,
  time = seasonalData_subsetQB_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
  mtry = 2,
  ntree = 500,
  sto = "BM")
